\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-10-03 paper/main.tex
working on intro...

Jupyter (iPython) notebooks~\citep{KluyEtal16} have revolutionized
code sharing by providing a single filetype that combines text, code,
and figures.  For projects that depend on other software packages,
however, additional work is often required to enable the code to run
as its authors intended.  For example, a piece of software that
depends on version 0.8 of package \textit{X} may behave differently
(or may fail to run at all) if the user instead attempts to run the
code in an environment with version 0.9 of package \textit{X}.

>>>>

Code sharing is a core component of the open science movement
that has inspired a full ecosystem of tools and packages.  However,
sharing code, in and of itself, does not guarantee that others will be
able to \textit{reproduce} the desired results.  For eample, research
code often requires installing other software packages that extends
the language's basic functionality.  Within the Python community,
external packages that are published in the most popular 
repositories~\citep{PIP, CONDA} are associated with version numbers
and tags that enable users to guarantee that they are installing
exactly the same code across different computing environments.
Despire that it is \textit{possible} to manually install the intended version numbers of
every dependency of a Python script or package, doing so may cause
conflicts within the user's computing environment that interferes with
the functionality of  \textit{other} code.

To facilitate code sharing, the Python community has developed a broad
set of approaches and tools (Fig.~\ref{fig:code-sharing}).  At one
extreme, simply publishing a set of Python scripts (.py files) may
enable others to use or gain insights into the relevant work.  Because
Python is installed by default on most modern operating systems, for
some projects this may be sufficient.  Another popular approach entails
creating JSON files, called Jupyter notebooks~\citep{KluyEtal16} that
comprise a mix of text, executable code, and embedded media.
Notebooks may call or import external scripts or libraries in order to
provide a more compact and readable experience for the users.  Each of
these systems (Python scripts and notebooks) provides a convenient
means of sharing code, with the caveat that they do not specify a
complete computing environment.  Therefore the functionality of code
shared using these systems cannot be guaranteed across different
computing environments.

At another extreme, virtual machines~citep{VM} provide a complete hardware-level
simulation of the desired system.  Virtual machines are typically
fully isolated from the user's system such that installing or running
software on a virtual machine does not impact the user's primary
operating system or computing environment.  Containers~\citep[e.g.,][]{Merk14,
  SINGULARITY} provide a similar ``isolated'' experience.  Although
containerized environments do not specify hardware-level operations,
they are typically packaged with a complete operating system, in
addition to a complete copy of Python and any relevant dependencies.
Virtual environments~\citep[e.g.,][]{vanRoEtal14} also provide a
computing environment that is largely separated from the user's main
environment.  They incorporate a copy of Python and the target
software's dependencies, but virtual environments do not specify or
reproducee a complete copy of the original operating system.  Each of
these systems (virtual machines, containers, and virtual environments)
guarantees (to differing degrees-- at the hardware-level, operating system
level, and Python environment, respectively) that the relevant code
will run similarly for different users.  However, each of these
systems also requires installing independent tools that
can be resource intensive or burdensome to install or configure.

We designed \texttt{davos} to occupy a ``sweet spot'' between these
extremes.  \texttt{davos} is a notebook-installable package that adds
functionality to the original.  Like standard Jupyter notebooks,
\texttt{davos}-enhanced notebooks allows researchers to include text,
executable code, and media within a single file.  No further setup or
installation is required, beyond what is needed to run standard
Jupyter notebooks.  And like virtual environments \texttt{davos}
provides a convenient mechanism for fully specifying (and installing,
as needed) a complete set of Python dependencies, including the
relevant version numbers.
